Actuator line modeling of vertical-axis turbines

### 2018-05-05 reviews/reviewer1-response.md

>Fig. 11. The results of power coefficient from the second case show that the model is quite case dependent. The results show that the model is not able to represent the physics about this turbine and thus the actual accuracy of the ALM cannot be verified without any additional figure that shows how the flow field is. Can you provide transversal plane contours showing the actual flow field generated by the ALM? Could this inaccuracy be again due to the k-epsilon model?

We have added a snapshot of the internal flow field and force vectors to give
and idea of the level of detail, or lack thereof, regarding local flow near
the blades. Unfortunately there is no mechanism for the ALM to generate flow
features like the pair of counter rotating vortices induced by dynamic stall.
This is a compromise inherent to the ALM, which we have clarified. It is
advantageous when attempting to simulate array-scale flows, but not so much
for individual device optimization, despite our validation efforts focusing
on the latter.

>Section 9.4. The presented LES results are quite disappointing due to the large differences against the RANS results which were very good. The reviewer has few concerns about this section:
>
Can you specify more details about the number of processors used? It is true that LES demands more computational resources than RANS but with the meshes used for the ALM it should be affordable to run LES.

We ran the LES simulations on 4 processors, a note that has been added to the
paper. It was still relatively affordable compared to the blade-resolved RANS.
We have noted a 2 order of magnitude drop in computational cost as well.

>Overall LES results are poor. The predicted Cp dropped to 20% which seems to be derived from the resolution of the flow field instead of the ALM itself. While a good agreement is found for the velocity wake as seen in Fig. 13 and 15(left), the turbulent kinetic energy results show an unexpected over damping of the velocity fluctuations. The use of LES is herein motivated to achieve a better resolution of the turbulent fluctuations by explicitly resolving the large-scale flow structures, and in the present manuscript the TKE has dropped dramatically. Fig. 14 and 15 (right) show a lack of flow unsteadiness attributed by the authors to the influence of the SGS model. The spatial discretisation is not described and it may be possible that depending on it, the scheme introduced an extra damping typically used in implicit LES wherein no SGS is used. Nevertheless, the reviewer would recommend to double-check the kinetic energy results as is quite confusing that the velocity wake prediction is greatly improved while the velocity fluctuations are almost gone.
>
>For further validation, the authors may consider to plot RANS vs LES at closer distance to the turbine, e.g. x/D=0.5 and study how the turbulent kinetic energy decay is in both models, or a transversal plane showing the generation of the vortices.

It is not likely that the numerical schemes are dissipating the energy, since
all 2nd order schemes are used. We went back and looked at contours of the
resolved and SGS TKE, and interestingly, the total TKE is significantly
increased after the transition in the grid resolution in the near-wake.
This does imply that the vortices created by the actuator lines do remain
artificially stable, and the grid transition helps break them down and generate
resaonably levels of TKE. This would not be much of an issue in array
simulations with spacings greater than a few diameters. A note has been added
to the manuscript. We considered adding the contour plots, but the paper is
starting to get quite long. We can include them if they will contribute to
the understanding, however.

### 2018-05-05 paper.tex
Finish addressing reviewer 1 comments

Contours of turbulence kinetic energy sampled at $x/D=1$ from the large eddy
simulation are plotted in Figure~\ref{fig:RVAT-ALM-LES-kcont}. Compared with
RANS, LES is more able to predict the turbulence generated by the blade tip
vortex shedding and dynamic stall effects, though the overall level of
unsteadiness generated was much lower, especially on the $+y$ side of the rotor.
This is likely a consequence of the SGS modeling, where the vortical structures
generated by the blades remain stable further downstream. Similar effects were
seen in \cite{Martinez-Tossas2015a, Shamsoddin2014}, where higher levels of the
Smagorinsky coefficient delayed vortex breakdown and subsequent higher levels of
turbulence. Figure~\ref{RVAT-ALM-LES-vorticity} shows evidence of these effects,
where the blade bound and tip vortices are still relatively coherent at $x/D=1$.

@@

Mean velocity profiles at the turbine center plane, plotted in
Figure~\ref{fig:RVAT-ALM-LES-profiles}, were predicted more accurately using LES
versus RANS, and rival those of the 3-D blade-resolved models. However,
turbulence kinetic energy profiles did not match as closely with experiments.
Though the qualitative shape was resolved better than that by the RANS ALM
simulation, notably the asymmetric peaks around $y/R = \pm 1$, the turbulence
generated in the large eddy simulation was approximately an order of magnitude
too low. Once again this was likely a result of the subgrid-scale modeling and
its effect on the stability of vortex structures, or its over-dissipative nature
removing energy from the flow before it is advected to the plane of interest.

>>>>

Contours of turbulence kinetic energy sampled at $x/D=1$ from the large eddy
simulation are plotted in Figure~\ref{fig:RVAT-ALM-LES-kcont}. Compared with
RANS, LES is more able to predict the turbulence generated by the blade tip
vortex shedding and dynamic stall effects, though the total turbulence kinetic
energy was lower, especially on the $+y$ side of the rotor.
This is likely a consequence of the SGS modeling, where the vortical structures
generated by the blades remain stable further downstream. Similar effects were
seen in \cite{Martinez-Tossas2015a, Shamsoddin2014}, where higher levels of the
Smagorinsky coefficient delayed vortex breakdown and subsequent higher levels of
turbulence. Figure~\ref{RVAT-ALM-LES-vorticity} shows evidence of these effects,
where the blade bound and tip vortices are still relatively coherent at $x/D=1$.

@@

Mean velocity profiles at the turbine center plane, plotted in
Figure~\ref{fig:RVAT-ALM-LES-profiles}, were predicted more accurately using LES
versus RANS, and rival those of the 3-D blade-resolved models. However,
turbulence kinetic energy profiles did not match as closely with experiments.
Though the qualitative shape was resolved better than that by the RANS ALM
simulation, notably the asymmetric peaks around $y/R = \pm 1$, the turbulence
generated in the large eddy simulation was approximately an order of magnitude
too low.
Note that $k$ (both resolved and SGS) does increase significantly downstream
after the transition in grid resolution, which implies the grid can help
compensate for the artificial stability of the vortex structures caused by
the Smagorinsky model.
It is hypothesized that the turbulence kinetic energy would better match
experimental data taken further downstream, meaning the model should perform
well for array simulations with spacing as close as $2D$.
